---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The script will load raw activity data and remove incomolete cases from the dataset.
The script downloads the data directly from the web and unzips it for use in a temporary location

```{r load data}
## Reads 

tFile <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", tFile)
fList <- unzip(tFile)
rawdataset <- read.csv(fList[1])
dataset <- rawdataset[complete.cases(rawdataset[1:3]),]

##read.csv(file.path(tDir,"activity.csv"))



```
## What is mean total number of steps taken per day?
Data for average steps per day.
The histogram shows the frequency with which particular numbers of steps occur on multiple days. The highest frequency is the modal number of steps.
```{r calculate mean number of steps}
library(dplyr)
dailytotals <- summarise(group_by(dataset, date), sum(steps))
hist(dailytotals$`sum(steps)`)
```
The mean and median numbers of steps follow:
```{r}
mean(dailytotals$`sum(steps)`)
median(dailytotals$`sum(steps)`)
```
## What is the average daily activity pattern?
```{r}

patterndata <- summarise(group_by(dataset, interval), mean(steps))
plot(patterndata$interval,patterndata$`mean(steps)`, type="l", xlab="5 minute interval", ylab="average steps taken")
```
## Imputing missing values
The total number of rows with missing values is:
```{r}
length(rawdataset[,1]) - length(dataset[,1]) 
```

This may have a significant impact on calculations, so 

```{r}
impute <- function(val, interval) {
  if (is.na(val)) {
    patterndata[patterndata$interval == 5,]$`mean(steps)`
  } else {
    val
  }
}
imputeddata <- rawdataset
for(i in 1:length(imputeddata[,1])) {
  imputeddata[i,1] <- impute(imputeddata[i,1],imputeddata[i,3])
}
imputeddata <- transmute(imputeddata, steps = impute(steps,interval))
## Are there differences in activity patterns between weekdays and weekends?
